\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-02-07 2-the-cubble-pkg.Rmd
go through all the fig captions

```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data. (1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "40%"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
```

@@

```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
```

>>>>

```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure of cube or hypercube.", fig.align='center', out.height = "40%"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
```

@@

```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the nested and long form in cubble. The nested form defines each site in a row and nests the time varying variables into a single column \\code{ts}. The long form cubble uses \\code{id} and \\code{t} to identify each row and store the time invariant variables as an attribute, \\code{spatial}."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
```

### 2022-02-07 3-advanced.Rmd
go through all the fig captions

```{r illu-hier, echo = FALSE, fig.align="center",out.height="40%", out.width = "100%", fig.cap = "Relationship between the station and cluster level cubble in both long and nested form."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
```

@@

```{r illu-matching, out.width="100%", fig.cap="An illustration of temporal matching in cubble. Three highest peaks are identified in each series and intervals are constructed on series A. Two peaks in series a fall into the intervals and hence the two series are considered to have 2 matches."}
knitr::include_graphics("figures/illu-matching.png")
```

@@

```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble. The line plot and map is made separately with the long and nested cubble. When a station is selected on the map (left), the corresponding row in the nested cubble will be activated. This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right)."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
```

>>>>

```{r illu-hier, echo = FALSE, fig.align="center",out.height="40%", out.width = "100%", fig.cap = "An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data. \\code{switch\\_key()} changes the station level cubble into a cluster level cubble and both can be stretched into the long form."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
```

@@

```{r illu-matching, out.width="100%", fig.cap="An illustration of temporal matching in cubble. Three highest peaks are identified in each series and intervals are constructed on series \\code{A}. Two peaks in series \\code{a} fall into the intervals and hence the two series are considered to have 2 matches."}
knitr::include_graphics("figures/illu-matching.png")
```

@@

```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble. The line plot and the map is made separately with the long and nested cubble. When a station is selected on the map (left), the corresponding row in the nested cubble will be activated. This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right)."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
```

### 2022-02-07 4-1-basic.Rmd
go through all the fig captions

```{r basic-manip, out.width="100%", out.height="70%", fig.cap="A glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations in Australia. On the top left corner is an insetted plot of station Cobar, which is highlighted in the black box."}
knitr::include_graphics("figures/basic-manip.png")
```

>>>>

```{r basic-manip, out.width="100%", out.height="70%", fig.cap="A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia. On the top left corner is an insetted plot of station Cobar highlighted in the black box."}
knitr::include_graphics("figures/basic-manip.png")
```

### 2022-02-07 4-2-agg.Rmd
go through all the fig captions

This example shows how to organise data at both level with \code{switch_key()}. The dataset used is \code{weatherdata::climate_full}, which records daily precipitation and maximum/ minimum temperature for \code{r nrow(weatherdata::climate_full)} stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used here to create 20 clusters. This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.

@@

```{r basic-agg, out.width="100%", fig.cap="Profile of aggregated precipitation from 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude of precipitation. Subplot B shows the station membership of each cluster"}
knitr::include_graphics("figures/basic-agg.png")
```

>>>>

This example shows how to organise data at both level with \code{switch_key()}. The dataset used is \code{weatherdata::climate_full}, which records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used here to create 20 clusters. This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.

@@

```{r basic-agg, out.width="100%", fig.cap="Profile of aggregated precipitation from 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster."}
knitr::include_graphics("figures/basic-agg.png")
```

### 2022-02-07 4-3-water-level.Rmd
go through all the fig captions

```{r matching-map, out.width="100%", fig.cap="Location of climate station and river gauges in Victoria, Australia."}
knitr::include_graphics("figures/matching-map.png")
```

@@

```{r matching, out.width="100%", fig.cap="Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The increases in the precipitation is reflected well by the water level."}
knitr::include_graphics("figures/matching.png")
```

>>>>

```{r matching-map, out.width="100%", fig.cap="Location of weather stations and river gauges in Victoria, Australia."}
knitr::include_graphics("figures/matching-map.png")
```

@@

```{r matching, out.width="100%", fig.cap="Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The increases in the precipitation is reflected by the water level."}
knitr::include_graphics("figures/matching.png")
```

### 2022-02-07 4-4-netcdf.Rmd
go through all the fig captions

NetCDF (Network Common Data Form) is a data format commonly used in climatology community to deliver global mapping of atmosphere, ocean, and land. It has two major components: dimension to define the spatiotemporal grid (longitude, latitude, and time) and variable which populates the defined grid. Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise this data format. Current packages on manipulating NetCDF data in R include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].

@@

```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row of Figure 19 in  Hersbach (2020)."}
knitr::include_graphics(here::here("figures/netcdf.png"))
```

>>>>

NetCDF (Network Common Data Form) is a data format commonly used in climatology community to deliver global mapping of atmosphere, ocean, and land. It has two major components: dimension to define the spatio-temporal grid (longitude, latitude, and time) and variable which populates the defined grid. Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise this data format. Current packages on manipulating NetCDF data in R include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].

@@

```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020)."}
knitr::include_graphics(here::here("figures/netcdf.png"))
```

### 2022-02-07 4-5-interactive.Rmd
add some sentences on data cube and fix \code{...} rendering
go through all the fig captions

With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns such as trend or seasonality in the time series. Combining this two types of plot with interactivity let users to link between points on the map and their corresponding time series and explore the spatial and temporal dimension of the data simultaneously. Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk{.

@@

```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display. Each row is a screen dump of the process. The top row shows all locations and all temperature profiles. Selecting a location with high variance on the map produces the plot in the second row. The maximum nad minimum temperature is shown using a ribbon. The bottom row first selects the lowest temperature in Auguest in the seasonal display. A location in the Tasmania Island is then selected and surprisingly, it is actually warmer than the Thredbo AWS.", fig.show='hold'}
knitr::include_graphics(here::here("figures/linking.png"))
knitr::include_graphics(here::here("figures/linking-north.png"))
knitr::include_graphics(here::here("figures/linking-lower.png"))
#knitr::include_graphics(here::here("figures/linking-south.png"))
```

@@

```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia. Upon clicking a single station on the leaftlet up, the temperature band will be shown as a subplot in the popup box.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup-mid.png")) 
```

>>>>

With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns such as trend or seasonality in the time series. Combining this two types of plot with interactivity let users to link between points on the map and their corresponding time series and explore the spatial and temporal dimension of the data simultaneously. Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.

@@

```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display. Each row is a screen dump of the process. The top row shows all locations and all temperature profiles. Selecting a location with high variance on the map produces the plot in the second row. The maximum nad minimum temperature is shown using a ribbon. The bottom row first selects the lowest temperature in Auguest in the seasonal display. A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.", fig.show='hold'}
knitr::include_graphics(here::here("figures/linking.png"))
knitr::include_graphics(here::here("figures/linking-north.png"))
knitr::include_graphics(here::here("figures/linking-lower.png"))
#knitr::include_graphics(here::here("figures/linking-south.png"))
```

@@

```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as FIgure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup-mid.png")) 
```

### 2022-02-07 5-conclusion.Rmd
all the  -> \code{CODE} as per jss  guideline
go through all the fig captions

@@

This work is funded by the Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61. The article is created using `knitr` [@knitr] and `rmarkdown` [@rmarkdown] in R. The source code for reproducing this paper can be found [here](https://github.com/huizezhang-sherry/paper-cubble).

@@

```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
```

>>>>

This paper describes the \proglang{R} package \pkg{cubble} for manipulating and visualising spatio-temporal data. A new data structure, that builds from the \code{rowwise_df} and \code{grouped_df} class in the tidyverse ecosystem, is proposed to connect the spatial and temporal variables in the spatio-temporal data. This design frees the data analysts from spending time on organising variables of different observational units. The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another pacakge to cluster stations. Further development and maintenance of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}. Another area for further development is to extend \pkg{cubble} to other spatial objects other than points.

@@

This work is funded by the Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61. The article is created using \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in R. The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.

@@

```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble. When a point on the time series is selected, the corresponding row in the long cubble will be activated. This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the smae id (middle). Both plots will be updated with the full line selected and the point highlighted on the map (right)."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
```

### 2022-02-07 main.tex
go through all the fig captions

\caption[Illustration of incoming data formats for spatio-temporal data]{Illustration of incoming data formats for spatio-temporal data. (1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.}\label{fig:illu-input}
\end{figure}
\end{CodeChunk}

@@

\caption[Illustration of nested and long cubble]{Illustration of nested and long cubble.}\label{fig:illu-cubble}
\end{figure}
\end{CodeChunk}

@@

\caption[Relationship between the station and cluster level cubble in both long and nested form]{Relationship between the station and cluster level cubble in both long and nested form.}\label{fig:illu-hier}
\end{figure}
\end{CodeChunk}

@@

\caption[An illustration of temporal matching in cubble]{An illustration of temporal matching in cubble. Three highest peaks are identified in each series and intervals are constructed on series A. Two peaks in series a fall into the intervals and hence the two series are considered to have 2 matches.}\label{fig:illu-matching}
\end{figure}
\end{CodeChunk}

@@

\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble. The line plot and map is made separately with the long and nested cubble. When a station is selected on the map (left), the corresponding row in the nested cubble will be activated. This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right).}\label{fig:illu-interactive}
\end{figure}
\end{CodeChunk}

@@

\caption[A glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations in Australia]{A glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations in Australia. On the top left corner is an insetted plot of station Cobar, which is highlighted in the black box.}\label{fig:basic-manip}
\end{figure}
\end{CodeChunk}

@@

This example shows how to organise data at both level with
\code{switch_key()}. The dataset used is
\code{weatherdata::climate_full}, which records daily precipitation and
maximum/ minimum temperature for
\code{r nrow(weatherdata::climate_full)} stations in Australia from 2016
to 2020. A simple kmean algorithm based on the distance matrix is used
here to create 20 clusters. This creates \code{station_nested} as a
station level nested cubble with a cluster column indicating the group
each station belongs to. More complex algorithms can be used for other
problem, as long as there is a mapping from each station to a cluster.

@@

\caption[Profile of aggregated precipitation from 639 weather stations in Australia]{Profile of aggregated precipitation from 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude of precipitation. Subplot B shows the station membership of each cluster}\label{fig:basic-agg}
\end{figure}
\end{CodeChunk}

@@

\caption[Location of climate station and river gauges in Victoria, Australia]{Location of climate station and river gauges in Victoria, Australia.}\label{fig:matching-map}
\end{figure}
\end{CodeChunk}

@@

\caption[Matched weather stations and river gauges on the map (A) and across time (B)]{Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The increases in the precipitation is reflected well by the water level.}\label{fig:matching}
\end{figure}
\end{CodeChunk}

@@

NetCDF (Network Common Data Form) is a data format commonly used in
climatology community to deliver global mapping of atmosphere, ocean,
and land. It has two major components: dimension to define the
spatiotemporal grid (longitude, latitude, and time) and variable which
populates the defined grid. Attributes are usually associated with
dimension and variable in the NetCDF format data and a
\href{http://cfconventions.org/}{metadata convention for climate and
forecast} has been designed to standardise this data format. Current
packages on manipulating NetCDF data in R include a high-level R
interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls
C-interface: \pkg{RNetCDF} \citep{rnetcdf, michna2013rnetcdf}, and a
tidyverse implementation: \pkg{tidync} \citep{tidync}.

@@

\caption[A reproduction of the second row of Figure 19 in  Hersbach (2020)]{A reproduction of the second row of Figure 19 in  Hersbach (2020).}\label{fig:netcdf}
\end{figure}
\end{CodeChunk}

@@

\caption[Exploring temperature variation using linking of a map and seasonal display]{Exploring temperature variation using linking of a map and seasonal display. Each row is a screen dump of the process. The top row shows all locations and all temperature profiles. Selecting a location with high variance on the map produces the plot in the second row. The maximum nad minimum temperature is shown using a ribbon. The bottom row first selects the lowest temperature in Auguest in the seasonal display. A location in the Tasmania Island is then selected and surprisingly, it is actually warmer than the Thredbo AWS.}\label{fig:interactive-linking}
\end{figure}
\end{CodeChunk}

@@

\caption[Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia]{Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia. Upon clicking a single station on the leaftlet up, the temperature band will be shown as a subplot in the popup box.}\label{fig:interactive-popup}
\end{figure}
\end{CodeChunk}

@@

This paper describes the \proglang{R} package \texttt{cubble} for
manipulating and visualising spatio-temporal data. A new data structure,
that builds from the \texttt{rowwise\_df} and \texttt{grouped\_df} class
in the tidyverse ecosystem, is proposed to connect the spatial and
temporal variables in the spatio-temporal data. This design frees the
data analysts from spending time on organising variables of different
observational units. The data structure is also flexible to the
techniques and packages analysts use to analyse the data, for example,
in the matching example in section 4.3, users are free to use algorithms
from another pacakge to cluster stations. Further development and
maintenance of the package involves responding to changes in the
tidyverse packages that \texttt{cubble} imports, in particular,
\texttt{tibble}, \texttt{tidyr}, and \texttt{dplyr}. Another area for
further development is to extend \texttt{cubble} to other spatial
objects other than points.

@@

\caption[demon interactivity]{demon interactivity}\label{fig:illu-interactive-2}
\end{figure}
\end{CodeChunk}

>>>>

\caption[An illustration of different spatio-temporal data format]{An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure of cube or hypercube.}\label{fig:illu-input}
\end{figure}
\end{CodeChunk}

@@

\caption{An illustration of the nested and long form in cubble. The nested form defines each site in a row and nests the time varying variables into a single column \code{ts}. The long form cubble uses \code{id} and \code{t} to identify each row and store the time invariant variables as an attribute, \code{spatial}.}\label{fig:illu-cubble}
\end{figure}
\end{CodeChunk}

@@

\caption[An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data]{An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data. \code{switch\_key()} changes the station level cubble into a cluster level cubble and both can be stretched into the long form.}\label{fig:illu-hier}
\end{figure}
\end{CodeChunk}

@@

\caption{An illustration of temporal matching in cubble. Three highest peaks are identified in each series and intervals are constructed on series \code{A}. Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have 2 matches.}\label{fig:illu-matching}
\end{figure}
\end{CodeChunk}

@@

\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble. The line plot and the map is made separately with the long and nested cubble. When a station is selected on the map (left), the corresponding row in the nested cubble will be activated. This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right).}\label{fig:illu-interactive}
\end{figure}
\end{CodeChunk}

@@

\caption[A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia]{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia. On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
\end{figure}
\end{CodeChunk}

@@

This example shows how to organise data at both level with
\code{switch_key()}. The dataset used is
\code{weatherdata::climate_full}, which records daily precipitation and
maximum/ minimum temperature for 639 stations in Australia from 2016 to
2020. A simple kmean algorithm based on the distance matrix is used here
to create 20 clusters. This creates \code{station_nested} as a station
level nested cubble with a cluster column indicating the group each
station belongs to. More complex algorithms can be used for other
problem, as long as there is a mapping from each station to a cluster.

@@

\caption[Profile of aggregated precipitation from 639 weather stations in Australia]{Profile of aggregated precipitation from 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster.}\label{fig:basic-agg}
\end{figure}
\end{CodeChunk}

@@

\caption[Location of weather stations and river gauges in Victoria, Australia]{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
\end{figure}
\end{CodeChunk}

@@

\caption[Matched weather stations and river gauges on the map (A) and across time (B)]{Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The increases in the precipitation is reflected by the water level.}\label{fig:matching}
\end{figure}
\end{CodeChunk}

@@

NetCDF (Network Common Data Form) is a data format commonly used in
climatology community to deliver global mapping of atmosphere, ocean,
and land. It has two major components: dimension to define the
spatio-temporal grid (longitude, latitude, and time) and variable which
populates the defined grid. Attributes are usually associated with
dimension and variable in the NetCDF format data and a
\href{http://cfconventions.org/}{metadata convention for climate and
forecast} has been designed to standardise this data format. Current
packages on manipulating NetCDF data in R include a high-level R
interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls
C-interface: \pkg{RNetCDF} \citep{rnetcdf, michna2013rnetcdf}, and a
tidyverse implementation: \pkg{tidync} \citep{tidync}.

@@

\caption[A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020)]{A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020).}\label{fig:netcdf}
\end{figure}
\end{CodeChunk}

@@

\caption[Exploring temperature variation using linking of a map and seasonal display]{Exploring temperature variation using linking of a map and seasonal display. Each row is a screen dump of the process. The top row shows all locations and all temperature profiles. Selecting a location with high variance on the map produces the plot in the second row. The maximum nad minimum temperature is shown using a ribbon. The bottom row first selects the lowest temperature in Auguest in the seasonal display. A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
\end{figure}
\end{CodeChunk}

@@

\caption[Same as FIgure 11 with thetemperature variation shown as a popup in the leaflet map]{Same as FIgure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
\end{figure}
\end{CodeChunk}

@@

This paper describes the \proglang{R} package \pkg{cubble} for
manipulating and visualising spatio-temporal data. A new data structure,
that builds from the \code{rowwise_df} and \code{grouped_df} class in
the tidyverse ecosystem, is proposed to connect the spatial and temporal
variables in the spatio-temporal data. This design frees the data
analysts from spending time on organising variables of different
observational units. The data structure is also flexible to the
techniques and packages analysts use to analyse the data, for example,
in the matching example in section 4.3, users are free to use algorithms
from another pacakge to cluster stations. Further development and
maintenance of the package involves responding to changes in the
tidyverse packages that \pkg{cubble} imports, in particular,
\pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}. Another area for further
development is to extend \pkg{cubble} to other spatial objects other
than points.

@@

\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble. When a point on the time series is selected, the corresponding row in the long cubble will be activated. This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the smae id (middle). Both plots will be updated with the full line selected and the point highlighted on the map (right).}\label{fig:illu-interactive-2}
\end{figure}
\end{CodeChunk}
